---
---

## Specialized Packages

The third tier of access to online data is much preferred, if it
exists: a dedicated package in your programming language's repository,
[CRAN](http://cran.r-project.org) or [PyPI](http://pypi.python.org).

- Additional guidance on query parameters
- Returns data in native formats
- Handles all "encoding" problems

===

The [tidycensus](){:.rlib} package, developed by Kyle Walker,
streamlines access to the API and is integrated with [tidyverse](){:.rlib} packages.

To repeat the exercise below at home, request an API key at
https://api.census.gov/data/key_signup.html, and store it in a file named `census_api_key.R`
in your working directory. The file should contain the line 
`Sys.setenv(CENSUS_API_KEY = 'your many digit key')`. This creates a hidden
environment variable containing the key.
{:.notes}


```{r, handout = 0, message = FALSE}
library(tidycensus)
source('census_api_key.R')
```

===

Compared to using the API directly via the [requests](){:.pylib} package:

**Pros**
- More concise code, quicker development
- Package documentation (if present) is usually more user-friendly than API documentaion.
- May allow seamless update if API changes

**Cons**
- No guarantee of updates
- Possibly limited in scope

===

Query the Census ACS5 survey (American Community Survey) for the variable `B19001_001E` (median annual household income,
in dollars) and each entity's `NAME`.

The American Community Survey (ACS) is a yearly survey that provides detailed population
and housing information at fine geographic scale across the United States. Much of the 
[tidycensus](){:.rlib} package is dedicated to accessing the ACS data. ACS5 refers to a five-year
average of the annual surveys.
{:.notes}

```{r, handout = 0}
variables <- c('NAME', 'B19013_001E')
```

===

This code pulls the variables `NAME` and `B19001_001E` from all counties
in Maryland. The [tidycensus](){:.rlib} package 
converts the JSON string into a data frame. (No need to check headers.) 

```{r, handout = 0}
county_income <- get_acs(geography = 'county',
                         variables = variables,
                         state = 'MD',
                         year = 2018,
                         geometry = TRUE)

head(county_income)
```

===

We can use [dplyr](){.rlib} to manipulate the output, and [ggplot2](){.rlib} to visualize the data.
Because we set `geometry = TRUE`, [tidycensus](){.rlib} even includes spatial information in its
output that we can use to create maps!

```{r, handout = 0}
ggplot(county_income) + 
  geom_sf(aes(fill = estimate), color = NA) + 
  coord_sf() + 
  theme_minimal() + 
  scale_fill_viridis_c()
```

