---
---

## Specialized Packages

The third tier of access to online data is much preferred, if it
exists: a dedicated package in your programming language's repository
[CRAN](http://cran.r-project.org)) or ([PyPI](http://pypi.python.org).

- Additional guidance on query parameters
- Returns data in native formats
- Handles all "encoding" problems

===

The [censusapi](){:.rlib} package, developed by Hannah Recht,
is a user-contributed suite of tools that streamline access to the API.

To repeat the exercise below at home, request an API key at
https://api.census.gov/data/key_signup.html, and store it in a file named `census_api_key.R`
in your working directory. The file should contain the line 
`Sys.setenv(CENSUS_KEY = 'your many digit key')`. This creates a hidden
environment variable containing the key.
{:.notes}


```{r, handout = 0}
library(censusapi)
source('census_api_key.R')
```

===

Compared to using the API directly via the [requests](){:.pylib} package:

**Pros**
- More concise code, quicker development
- Package documentation (if present) is usually more user-friendly than API documentaion.
- May allow seamless update if API changes

**Cons**
- No guarantee of updates
- Possibly limited in scope

===

Query the Census ACS5 survey (American Community Survey) for the variable `B19001_001E` (median annual household income,
in dollars) and each entity's `NAME`.

The American Community Survey (ACS) is a yearly survey that provides detailed population
and housing information at fine geographic scale across the United States. Much of the 
[censusapi](){:.rlib} package is dedicated to accessing the ACS data. ACS5 refers to a five-year
average of the annual surveys.
{:.notes}

```{r, handout = 0}
variables <- c('NAME', 'B19013_001E')
```

===

This code pulls the variables `NAME` and `B19001_001E` from all census tracts and all
counties in the state with ID `24` (Maryland). The [censusapi](){:.rlib} package 
converts the JSON string into a data frame. (No need to check headers.) 

```{r, handout = 0}
maryland_income_by_tract <- getCensus(name = 'acs/acs5', 
                                      vintage = 2017, 
                                      vars = variables, 
                                      region = 'tract:*', 
                                      regionin = 'state:24+county:*')

head(maryland_income_by_tract)
```

===

We can use our `dplyr` tools to manipulate the output of `getCensus`, 
and `ggplot2` to create boxplots showing the income distribution among
census tracts within each county in Maryland.

```{r, handout = 0}
library(ggplot2)

maryland_income_by_tract <- maryland_income_by_tract %>%
  rename(median_household_income = B19013_001E) %>%
  filter(median_household_income > 0) 

ggplot(maryland_income_by_tract, aes(x = county, y = median_household_income)) +
  geom_boxplot()

```

