---
---

## Paging & Stashing

A common strategy that web service providers take to balance their
load is to limit the number of records a single API request can
return. The user ends up having to flip through "pages" with the API,
handling the response content at each iteration. Options for stashing
data are:

1. Store it all in memory, write to file at the end.
1. Append each response to a file, writing frequently.
1. Offload these decisions to database management software.

The [data.gov](https://www.data.gov) API provides a case in point. 
Data.gov is a service provided by the U.S. federal government to make data available
from across many government agencies. It hosts a catalog of raw data and of many other
APIs from across government.
Among the APIs catalogued by data.gov is the [FoodData Central](https://fdc.nal.usda.gov/) API.
The U.S. Department of Agriculture maintains a data system of nutrition information 
for thousands of foods. 
We might be interested in the relative nutrient content of different fruits.
{:.notes}

To repeat the exercise below at home, request an API key at
https://api.data.gov/signup/, and store it in a file named `datagov_api_key.R`
in your working directory. The file should contain the line 
`Sys.setenv(DATAGOV_KEY = 'your many digit key')`.
{:.notes}

===

Load the `DATAGOV_KEY` variable as an environment variable by importing it from the file you saved it in.

```{r}
source('datagov_api_key.R')
```

===

Run an API query for all foods with `"fruit"` in their name.

```{r, handout = 0}
api <- 'https://api.nal.usda.gov/fdc/v1/'
path <- 'foods/search'

query_params <- list('api_key' = Sys.getenv('DATAGOV_KEY'),
                     'query' = 'fruit')

doc <- GET(paste0(api, path), query = query_params) %>%
  content(as = 'parsed')
```

===

Extract data from the returned JSON object, which gets mapped to an
R list called `doc`.
First inspect the names of the list elements.

```{r}
names(doc)
```

===

We can print the value of `doc$totalHits` to see
how many foods matched our search term, `"fruit"`.

```{r}
doc$totalHits
```

===

The purported claimed number of results is much larger than the length
of the `foods` array contained in this response. The query returned only the
first page, with 50 items.

```{r}
length(doc$foods)
```

===

Continue to inspect the returned object. Extract one element from the list
of `foods` and view its description.

```{r}
fruit <- doc$foods[[1]]
fruit$description
```

===

Use the `purrr` list manipulation package to view the first ten 
nutrients in the food item and their amounts.

```{r, handout = 0}
map_dfr(fruit$foodNutrients, 
        ~ data.frame(name = .$nutrientName, value = .$value)) %>%
  head(10)
```

===

The `DBI` and `RSQLite` packages together allow R to connect to a 
database-in-a-file. If the `fruits.sqlite` file does not exist
in your working directory already when you try to connect,
`dbConnect()` will create it.

```{r, handout = 0}
library(DBI) 
library(RSQLite)

fruit_db <- dbConnect(SQLite(), 'fruits.sqlite') 
```

===

Add a new `pageSize` parameter to request `100` documents per page.

```{r, handout = 0}
query_params$pageSize <- 100
```

===

In each request (each iteration through the loop), 
advance the query parameter `pageNumber` by one. 
The first record retrieved will be `pageNumber * pageSize`. 

We use some `tidyr` and `dplyr` manipulations to
extract the ID number, name, and the amount of sugar from each
of the foods in the page of results returned by the query. 
This long manipulation is necessary because R does not easily handle the
nested list structures that APIs return. If we were using
a specialized API R package, typically it would handle this data wrangling 
for us.
{:.notes}

Insert the fruits (the three-column data frame `values`) 
in bulk to the database with `dbWriteTable()`.


```{r, handout = 0}
for (i in 1:10) {
  # Advance page and query
  query_params$pageNumber <- i
  response <- GET(paste0(api, path), query = query_params) 
  page <- content(response, as = 'parsed')
  fruits <- page$foods
  
  # Convert nested list to data frame
  values <- tibble(fruit = fruits) %>%
    unnest_wider(fruit) %>%
    unnest_longer(foodNutrients) %>%
    unnest_wider(foodNutrients) %>%
    filter(grepl('Sugars, total', nutrientName)) %>%
    select(fdcId, description, value) %>%
    setNames(c('foodID', 'name', 'sugar'))

  # Stash in database
  dbWriteTable(fruit_db, name = 'Food', value = values, append = TRUE)

}
```

===

View the records in the database by reading
everything we have so far into a data frame.

```{r, handout = 0}
fruit_sugar_content <- dbReadTable(fruit_db, name = 'Food')
head(fruit_sugar_content, 10)
```

===

Don't forget to disconnect from your database!

```{r, handout = 0}
dbDisconnect(fruit_db)
```


